Hazelnut: A Bidirectionally Typed \\ Structure Editor Calculus

### 2016-06-25 intro.tex
added citations

%Although programs can be represented as text, most text does not correspond to a syntactically well-formed and semantically well-defined program. 
Most programming languages define a textual syntax. This allows programmers to use text editors and other standard text-based tools to construct and manipulate programs. While this is of substantial practical utility, it also introduces some fundamental complexity into the programming process.

Cognitively, it requires that programmers understand the subtleties of text parsing (e.g. the relative precedence of the various forms.) This can be particularly challenging for novice and end-user programmers.\todo{citation?}

It also complicates matters for tool designers because they might be confronted with text that does not correspond to a well-formed, meaningful program\todo{mention YoungSeok's data here}. This may be because the programmer is in the midst of a sequence of 
edit actions that leaves the text temporarily malformed or ill-typed, or because the programmer has made a mistake. The language definition is silent about these situations, so it is difficult for tools to help programmers determine and execute a corrective course of action (e.g. by providing \todo{cite study on benefits of syntax highlighting}syntax highlighting and semantics-aware code completion services\todo{cite something?}.)

>>>>

%Although programs can be represented as text, most text does not correspond to a syntactically well-formed and semantically well-defined program.

Every day, programmers use text-based tools to construct and manipulate programs.  
These programs are written in languages which are defined by a textual syntax. 
While textual syntax has proved to have practical utility, it also introduces some fundamental complexity into the programming process.

%Doesn't really start out with a bang, IMHO. I like to start papers out with a "big" problem.
%Most programming languages define a textual syntax. This allows programmers to use text editors and other standard text-based tools to construct and manipulate programs. While this is of substantial practical utility, it also introduces some fundamental complexity into the programming process.

Cognitively, it requires that programmers understand the subtleties of text parsing (e.g. the relative precedence of the various forms.) This can be particularly challenging for novice and end-user programmers.
One study~\cite{Altadmri:2015:MCI:2676723.2677258} of novice programmer error found that the most common of all errors was an syntactic error, specifically unbalanced parenthesis.
% Too much detail?

It also complicates matters for tool designers because they might be confronted with text that does not correspond to a well-formed, meaningful program
By analyzing the same data set used in~\cite{6883030}, which contains 1,460 hours of fine-grained code editing logs collected from 21 programmers, we found that 44.2\% of the code edits resulted in syntactically malformed state of the source code.
%\todo{mention YoungSeok's data here}. 
This may be because the programmer is in the midst of a sequence of 
edit actions that leaves the text temporarily malformed or ill-typed, or because the programmer has made a mistake. The language definition is silent about these situations, so it is difficult for tools to help programmers determine and execute a corrective course of action (e.g. by providing \todo{cite study on benefits of syntax highlighting}syntax highlighting and semantics-aware code completion services\todo{cite something?}.)
